---
title: "DataProcessing"
author: "Erika Munshi"
date: "4/20/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Reading Data In
```{r}
getwd()
library(tidyverse)
library(lubridate)
ForestData <- read.csv("../RawData/TreeLoss.csv")
CropData <- read.csv("../RawData/CropData.csv")
PopData <- read.csv("../RawData/PopulationData.csv")
```

##Creating One Dataset
### First step is to create a new dataset to contain all three datasets
### Then isolate the yield and area data by crop 
### The next step is to select columns from the crop datasets and add them to the full dataset
### The last step is to add the population data from population dataset to the full dataset
### Finally we write a new csv with the full dataset

```{r}
#Create a full dataframe
FullData <- as.data.frame(ForestData)

#Isolate crop area variables
Cassava.Area <- 
  CropData %>%
  filter(Item == "Cassava" & Element=="Area harvested")
Cloves.Area <- 
  CropData %>%
  filter(Item == "Cloves" & Element=="Area harvested")
Cocoa.Area <- 
  CropData %>%
  filter(Item == "Cocoa, beans" & Element=="Area harvested")
Coffee.Area <- 
  CropData %>%
  filter(Item == "Coffee, green" & Element=="Area harvested")
Maize.Area <- 
  CropData %>%
  filter(Item == "Maize" & Element=="Area harvested")
Rice.Area <- 
  CropData %>%
  filter(Item == "Rice, paddy" & Element=="Area harvested")
SweetPotato.Area <- 
  CropData %>%
  filter(Item == "Sweet potatoes" & Element=="Area harvested")
Vanilla.Area <- 
  CropData %>%
  filter(Item == "Vanilla" & Element=="Area harvested")

#Isolate crop yield variables
Cassava.Yield <- 
  CropData %>%
  filter(Item == "Cassava" & Element=="Yield")
Cloves.Yield <- 
  CropData %>%
  filter(Item == "Cloves" & Element=="Yield")
Cocoa.Yield <- 
  CropData %>%
  filter(Item == "Cocoa, beans" & Element=="Yield")
Coffee.Yield <- 
  CropData %>%
  filter(Item == "Coffee, green" & Element=="Yield")
Maize.Yield <- 
  CropData %>%
  filter(Item == "Maize" & Element=="Yield")
Rice.Yield <- 
  CropData %>%
  filter(Item == "Rice, paddy" & Element=="Yield")
SweetPotato.Yield <- 
  CropData %>%
  filter(Item == "Sweet potatoes" & Element=="Yield")
Vanilla.Yield <- 
  CropData %>%
  filter(Item == "Vanilla" & Element=="Yield")

#Adding crop data to full dataframe         
FullData$CassavaArea <- Cassava.Area$Value
FullData$CassavaYield <- Cassava.Yield$Value
FullData$ClovesArea <- Cloves.Area$Value
FullData$ClovesYield <- Cloves.Yield$Value
FullData$CocoaArea <- Cocoa.Area$Value
FullData$CocoaYield <- Cocoa.Yield$Value
FullData$CoffeeArea <- Coffee.Area$Value
FullData$CoffeeYield <- Coffee.Yield$Value
FullData$MaizeArea <- Maize.Area$Value
FullData$MaizeYield <- Maize.Yield$Value
FullData$RiceArea <- Rice.Area$Value
FullData$RiceYield <- Rice.Yield$Value
FullData$SweetPotatoArea <- SweetPotato.Area$Value
FullData$SweetPotatoYield <- SweetPotato.Yield$Value
FullData$VanillaArea <- Vanilla.Area$Value
FullData$VanillaYield <- Vanilla.Yield$Value

#Adding Population Data
FullData$Population <- PopData$Value

#Writing a csv
write.csv(FullData, row.names = FALSE, file = "../ProcessedData/FullData.csv")

```